---
title: "Assignment1_Template"
format: html
execute:
   echo: false
---

```{r}
library(datasets)
library(knitr)
```



1) For the variable ‘Height’ in the ‘trees’ data set, complete the following tasks.

```{r}
# This will put the relevant data into a variable for you.
Assignment1_Data = trees$Height
```


A. Calculate the mean, variance, and standard deviation. (20 pts)

```{r}
# Put the variable below into the relevant functions to accomplish your task.
Assignment1_Data


```


B. Make a frequency distribution table. Make sure it includes the frequency, relative frequency, cumulative frequency and cumulative percentage. Do not exlude values which have a frequency of 0 (zero). Make sure your table is presented in APA style. Do not group the data into intervals. (60 points)

```{r}
# Create frequency table for Height, including every integer from the 
# minimum to the maximum even if some have zero frequency. This 
# leaves all options open for making a frequency table.
all_heights = min(trees$Height):max(trees$Height)
height = factor(trees$Height, levels = all_heights)
# This set of functions gives us the frequency for each level we just made and
# makes the table have the right format for our work.
freq = as.numeric(table(height))
# This gives us the total number of trees measured.
freq_total = sum(freq)
# Relative frequency is the frequency of a certain value relative to the total.
rel_freq = freq / freq_total
# This function adds in sequence the frequencies we have measured; i.e., the
# number of trees with the lowest height plus those with the next height...
cumu_f = cumsum(freq)
# The converts the cumulative fequency to a percentage of the total.
cumu_pct = cumu_f / freq_total * 100
# Build the table. In the data.frame function you write the name of the column,
# then the equal sign, then what you want to go in that column.
tab = data.frame(
  Height = all_heights, # Make a line in the table for each value
  f = freq,
  `Relative f` = round(rel_freq, 2), # Rounds to two decimal places
  `Cumulative f` = cumu_f,
  `Cumulative %` = round(cumu_pct),
  check.names = FALSE
)
# Add Total row
tab = rbind(tab, data.frame(
  Height = "Total",
  f = freq_total,
  `Relative f` = "",
  `Cumulative f` = "",
  `Cumulative %` = "",
  check.names = FALSE
))
# Print the table
kable(tab, row.names = FALSE)
```



C. Make a frequency distribution graph with the normal curve superimposed on it. Plot the scores on the x axis and the frequency on the y axis. Make sure your graph is presented in APA style. Do not group the data into intervals. (30 points)
```{r}
# Set a serifed font
par(family = "serif")

# Force every integer from min to max Height to appear
all_heights = min(Assignment1_Data):max(Assignment1_Data)
htab = table(factor(Assignment1_Data, levels = all_heights))
x_vals = as.numeric(names(htab))
y_freq = as.numeric(htab)

# Create the barplot (now with correct spacing)
bp = barplot(y_freq,
             space = 0,
             col = "white",
             ylab = "Frequency",
             xlab = "Height",
             cex.lab = 1.3,
             main = "",
             cex.names = 0.8,          # slightly smaller so labels fit
             cex.axis = 1,
             yaxs = "i",
             xaxs = "i",
             mgp = c(2.2, 0.5, 0),
             names.arg = x_vals)

#####################################################
##### Uncomment the next lines to add a normal curve.
# mu = mean(Assignment1_Data)
# sigma = sd(Assignment1_Data)
# n = length(Assignment1_Data)
# fine_x = seq(min(x_vals), max(x_vals), length.out = 300)
# dens_fine = dnorm(fine_x, mean = mu, sd = sigma) * n
# fine_bp = approx(x = x_vals, y = bp, xout = fine_x)$y
# lines(fine_bp, dens_fine, col = "black", lwd = 2)

# Restore default font
par(family = "")
```


D. Make a histogram of the frequency data. Looking at the histogram, describe the distribution as you have been instructed in class (not as ‘pretty,’ ‘fun,’ ‘strange,’ etc.). (15 points) 

```{r}
# Set a serifed font
par(family = "serif")

breaks = seq(from = min(Assignment1_Data),
             to   = max(Assignment1_Data) + 3,
             by   = 3)

labels = paste0(breaks[-length(breaks)], "-", breaks[-1] - 1)

A1_Data_Binned = cut(Assignment1_Data,
                     breaks = breaks,
                     labels = labels,
                     include.lowest = TRUE,
                     right = FALSE)

bp = barplot(table(A1_Data_Binned),
             space = 0,           
             col = "white",       
             ylab = "Frequency",
             xlab = "Heights",
             cex.lab = 1.3,       
             main = "",           
             cex.names = 1,       
             cex.axis = 1,        
             yaxs = "i",          
             xaxs = "i",
             mgp = c(2.2, 0.5, 0) 
)

# Restore default font
par(family = "")
```


E. Make a boxplot (10 points).
```{r}
boxplot(Assignment1_Data)
```




