---
title: "Assignment4"
format: html
execute:
   echo: false
   warning: false
   fig-dpi: 300
---


```{r}
library(tidyverse)
library(knitr)
library(car)
library(psych)
library(emmeans)
library(phia)
library(interactions)
```


```{r}
# Load your data
A4Data = read.table("A4Data.csv",
                    sep = ",",
                    header = T,
                    colClasses = c("factor",
                                   "factor",
                                   "factor",
                                   "integer"))
```

### Your data
Below is the data for a hypothetical factorial experiment assessing hypertension using mean arterial pressure (MAP) as a dependent variable. It is a 2 (dieting) X 2 (biofeedback) x 3 (drugs) between groups design (i.e., each participant provided one data point). Participants were assigned to a dieting condition (Diet: 1 = no, 2 = yes), a biofeedback condition (Bio: 1 = no, 2 = yes), and a drug condition (Drug: 1 = placebo, 2 = low dose, 3 = high dose). The outcome variable, MAP, was mean arterial pressure measured in mmHg.
```{r}
kable (A4Data,align = 'c', digits = 3)

vars = c("Diet", "Bio", "Drug")
a = NULL
for (v in vars) {
   for (lev in levels(A4Data[[v]])) {
      desc = describe(A4Data[A4Data[[v]] == lev, "MAP"])[c(2:4)]
      a = rbind(a, desc)
   }
}
rownames(a) = c(
   paste0("Diet", levels(A4Data$Diet)),
   paste0("Bio", levels(A4Data$Bio)),
   paste0("Drug", levels(A4Data$Drug))
)

kable(a, align = 'c', digits = 3)

```


```{r}

# "contr.sum" is the correct option for typle type III SS
options(contrasts = c("contr.sum", "contr.poly"))
# "contr.treatment" is a reference group type contrast closer to a type I or
# type II SS.
#options(contrasts = c("contr.treatment", "contr.poly"))

leveneTest(MAP ~ Diet * Bio * Drug, A4Data)
```

### Linear Model
```{r}
A4LM = lm(MAP ~ Diet * Bio * Drug, A4Data )

summary(A4LM)

```

# Omnibus
```{r}
a = Anova(A4LM, type = 3)

SS_error = a["Residuals", "Sum Sq"]

a$pes = a$`Sum Sq` / (a$`Sum Sq` + SS_error)
a["Residuals", "pes"] = NA          
a["(Intercept)", "pes"] = NA 
kable(a, align = 'c', digits = 3)
```

```{r}
# Post hoc adjustment type. Uncomment only one choice.
AdjustmentChoice = "none" # Least significant difference
# AdjustmentChoice = "bonferrroni"
# AdjustmentChoice = "tukey"
```

### Two-way comparisons

```{r}
# Cell means for Bio at each level of Diet
d = describeBy(MAP ~ Diet + Bio, data = A4Data, mat = TRUE, digits = 2)
names(d)[names(d) == "group1"] = "Diet"
names(d)[names(d) == "group2"] = "Bio"
rownames(d) = NULL
d[c(2,3,5:7,16)] %>%
  kable(caption = "Descriptives: Bio at each level of Diet", align = 'c', digits = 3)
```

```{r}
emmeans(A4LM, pairwise ~ Bio | Diet, adjust = AdjustmentChoice)$contrasts %>% 
  kable(caption = "Pairwise Bio at each level of Diet", align = 'c', digits = 3)
```

```{r}
# Cell means for Bio at each level of Drug
d = describeBy(MAP ~ Drug + Bio, data = A4Data, mat = TRUE, digits = 2)
names(d)[names(d) == "group1"] = "Drug"
names(d)[names(d) == "group2"] = "Bio"
rownames(d) = NULL
d[c(2,3,5:7,16)] %>%
  kable(caption = "Descriptives: Bio at each level of Drug", align = 'c', digits = 3)
```

```{r}
emmeans(A4LM, pairwise ~ Bio | Drug, adjust = AdjustmentChoice)$contrasts %>% 
  kable(caption = "Pairwise Bio at each level of Drug", align = 'c', digits = 3)
```

```{r}
# Cell means for Diet at each level of Drug
d = describeBy(MAP ~ Drug + Diet, data = A4Data, mat = TRUE, digits = 2)
names(d)[names(d) == "group1"] = "Drug"
names(d)[names(d) == "group2"] = "Diet"
rownames(d) = NULL
d[c(2,3,5:7,16)] %>%
  kable(caption = "Descriptives: Diet at each level of Drug", align = 'c', digits = 3)
```

```{r}
emmeans(A4LM, pairwise ~ Diet | Drug, adjust = AdjustmentChoice)$contrasts %>% 
  kable(caption = "Pairwise Diet at each level of Drug", align = 'c', digits = 3)
```

```{r}
# Cell means for Diet at each level of Bio
d = describeBy(MAP ~ Bio + Diet, data = A4Data, mat = TRUE, digits = 2)
names(d)[names(d) == "group1"] = "Bio"
names(d)[names(d) == "group2"] = "Diet"
rownames(d) = NULL
d[c(2,3,5:7,16)] %>%
  kable(caption = "Descriptives: Diet at each level of Bio", align = 'c', digits = 3)
```

```{r}
emmeans(A4LM, pairwise ~ Diet | Bio, adjust = AdjustmentChoice)$contrasts %>% 
  kable(caption = "Pairwise Diet at each level of Bio", align = 'c', digits = 3)
```

```{r}
# Cell means for Drug at each level of Diet
d = describeBy(MAP ~ Diet + Drug, data = A4Data, mat = TRUE, digits = 2)
names(d)[names(d) == "group1"] = "Diet"
names(d)[names(d) == "group2"] = "Drug"
rownames(d) = NULL
d[c(2,3,5:7,16)] %>%
  kable(caption = "Descriptives: Drug at each level of Diet", align = 'c', digits = 3)
```

### To justify t-tests within variables which have 3 or more levels you need a successful f-test.

```{r}
testInteractions(A4LM, 
                 fixed  = "Diet", 
                 across = "Drug") %>% kable(align = 'c', digits = 3)
```

```{r}
emmeans(A4LM, pairwise ~ Drug | Diet, adjust = AdjustmentChoice)$contrasts %>% 
  kable(caption = "Pairwise Drug at each level of Diet", align = 'c', digits = 3)
```

```{r}
# Cell means for Drug at each level of Bio
d = describeBy(MAP ~ Bio + Drug, data = A4Data, mat = TRUE, digits = 2)
names(d)[names(d) == "group1"] = "Bio"
names(d)[names(d) == "group2"] = "Drug"
rownames(d) = NULL
d[c(2,3,5:7,16)] %>%
  kable(caption = "Descriptives: Drug at each level of Bio", align = 'c', digits = 3)
```

### Again, to justify t-tests within variables which have 3 or more levels you need a successful f-test.

```{r}
testInteractions(A4LM, 
                 fixed  = "Bio", 
                 across = "Drug") %>% kable(align = 'c', digits = 3)
```

```{r}
emmeans(A4LM, pairwise ~ Drug | Bio, adjust = AdjustmentChoice)$contrasts %>% 
  kable(caption = "Pairwise Drug at each level of Bio", align = 'c', digits = 3)
```


### Three-way comparisons

```{r}
# Cell means for Diet at each Bio × Drug combination
d = describeBy(MAP ~ Bio + Drug + Diet, data = A4Data, mat = TRUE, digits = 2)
names(d)[names(d) == "group1"] = "Bio"
names(d)[names(d) == "group2"] = "Drug"
names(d)[names(d) == "group3"] = "Diet"
rownames(d) = NULL
d[c(2,3,4,6:8,17)] %>%
  kable(caption = "Descriptives: Diet at each Bio × Drug combination", align = 'c', digits = 3)
```

```{r}
emmeans(A4LM, pairwise ~ Diet | Bio * Drug, adjust = AdjustmentChoice)$contrasts %>%
  kable(caption = "Pairwise Diet at each Bio × Drug combination", align = 'c', digits = 3)
```

```{r}
# Cell means for Bio at each Diet × Drug combination
d = describeBy(MAP ~ Diet + Drug + Bio, data = A4Data, mat = TRUE, digits = 2)
names(d)[names(d) == "group1"] = "Diet"
names(d)[names(d) == "group2"] = "Drug"
names(d)[names(d) == "group3"] = "Bio"
rownames(d) = NULL
d[c(2,3,4,6:8,17)] %>%
  kable(caption = "Descriptives: Bio at each Diet × Drug combination", align = 'c', digits = 3)
```

```{r}
emmeans(A4LM, pairwise ~ Bio | Diet * Drug, adjust = "none")$contrasts %>%
  kable(caption = "Pairwise Bio at each Diet × Drug combination", align = 'c', digits = 3)
```

```{r}
# Cell means for Drug at each Diet × Bio combination
d = describeBy(MAP ~ Diet + Bio + Drug, data = A4Data, mat = TRUE, digits = 2)
names(d)[names(d) == "group1"] = "Diet"
names(d)[names(d) == "group2"] = "Bio"
names(d)[names(d) == "group3"] = "Drug"
rownames(d) = NULL
d[c(2,3,4,6:8,17)] %>%
  kable(caption = "Descriptives: Drug at each Diet × Bio combination", align = 'c', digits = 3)
```

### Again, to justify t-tests within variables which have 3 or more levels you need a successful f-test. In the first column, the first number is the Diet level the second is the Bio level.

```{r}
testInteractions(A4LM,
                 fixed = c("Diet", "Bio"),
                 across = "Drug") %>% 
  kable(align = 'c', digits = 3)
```

```{r}
emmeans(A4LM, pairwise ~ Drug | Diet * Bio, adjust = "none")$contrasts %>%
  kable(caption = "Pairwise Drug at each Diet × Bio combination", align = 'c', digits = 3)
```
