---
title: "Assignment4"
format: html
execute:
   echo: false
   warning: false
   fig-dpi: 300
---


```{r}
library(tidyverse)
library(knitr)
library(car)
library(psych)
library(emmeans)
library(phia)
library(interactions)
```


### Your data and analysis
If you correctly import your data you will see the output for a hypothetical factorial experiment assessing hypertension using mean arterial pressure (MAP), measured in mmHg, as a dependent variable. It is a 2 (dieting) X 2 (biofeedback) x 3 (drugs) between groups design (i.e., each participant provided one data point). Participants were assigned to a dieting condition (Diet: 1 = no, 2 = yes), a biofeedback condition (Bio: 1 = no, 2 = yes), and a drug condition (Drug: 1 = placebo, 2 = low dose, 3 = high dose).

```{r}
# Load your data and make some analysis choices in this code block.

# Put the full name of your data file in here, between the quotation marks, including the file extention.
NAMEOFYOURDATA = ""

# As listed in the description you have four variables. You need to tell the computer whether these are categorical or continuous. Your data is organized thusly:
# Column1- Drug
# Column2- Bio
# Column3- Diet
# Column4- MAP

# You can tell the computer your variable is categorical by puting the word factor between the quotation marks and continuous by putting the word integer (later this will be numeric).
Column1 = ""
Column2 = ""
Column3 = ""
Column4 = ""


###########################
### Your analysis choices.

### If you would like a test of homogeneity of variance for your data, set this variable to TRUE.
HOVTest = FALSE

# Choose the type of follow-up tests that you will use and set the comment marks correctly. Your choices are as follows.

# none - This will perform the 'Multiple t-test' method without correction.
# bonferroni - This will adjust the p-value in your 'Multiple t-test' follow-ups with a BF correction.
# tukey - This will perform a Tukey post hoc test instead of the 'Multiple tets' method.

AdjustmentChoice = "none"
# AdjustmentChoice = "bonferroni"
# AdjustmentChoice = "tukey"


```

### A) Render this mark down file and look at the omnibus output. Use it to manually calculate the effect size for each of your main effects and interactions. The 'Effect Sizes' table below the omnibus will count as your response.

```{r}
# Replace each 1 below with the partial eta-squared values you calculate from the Omnibus table. Don't change the names of the variables.
# Main effects
Diet_Peta = 1
Bio_Peta = 1
Drug_Peta = 1

# Two-way interactions
Diet_Bio_Peta = 1
Diet_Drug_Peta = 1
Bio_Drug_Peta = 1

# Three-way interaction
Diet_Bio_Drug_Peta = 1

```

### B) Write a results section (100 points). Note there is a results section example in your textbook (page 443).


---

---

### Raw data
```{r}


A4Data = read.table(NAMEOFYOURDATA,
                    sep = ",",
                    header = T,
                    colClasses = c(Column1,
                                   Column2,
                                   Column3,
                                   Column4))


kable (A4Data,align = 'c', digits = 3)
```

---

### Descriptives
```{r}
vars = c("Diet", "Bio", "Drug")
a = NULL
for (v in vars) {
   for (lev in levels(A4Data[[v]])) {
      desc = describe(A4Data[A4Data[[v]] == lev, "MAP"])[c(2:4)]
      a = rbind(a, desc)
   }
}
rownames(a) = c(
   paste0("Diet", levels(A4Data$Diet)),
   paste0("Bio", levels(A4Data$Bio)),
   paste0("Drug", levels(A4Data$Drug))
)

kable(a, align = 'c', digits = 3)

```

---

```{r}

# "contr.sum" is the correct option for typle type III SS
options(contrasts = c("contr.sum", "contr.poly"))
# "contr.treatment" is a reference group type contrast closer to a type I or
# type II SS.
#options(contrasts = c("contr.treatment", "contr.poly"))

if (HOVTest == T)
{
leveneTest(MAP ~ Diet * Bio * Drug, A4Data)
}
   
```

### Linear Model
```{r}
A4LM = lm(MAP ~ Diet * Bio * Drug, A4Data )

summary(A4LM)

```

# Omnibus

```{r}
a = Anova(A4LM, type = 3)

# SS_error = a["Residuals", "Sum Sq"]
# 
# a$pes = a$`Sum Sq` / (a$`Sum Sq` + SS_error)
# a["Residuals", "pes"] = NA
# a["(Intercept)", "pes"] = NA
kable(a, align = 'c', digits = 3)
```

### Effect Sizes 

```{r}
EffectSizes = data.frame(
  Term = c("Diet", "Bio", "Drug", 
           "Diet:Bio", "Diet:Drug", "Bio:Drug", 
           "Diet:Bio:Drug"),
  Partial_Eta_Squared = c(Diet_Peta, Bio_Peta, Drug_Peta, 
                          Diet_Bio_Peta, Diet_Drug_Peta, Bio_Drug_Peta, 
                          Diet_Bio_Drug_Peta)
)

kable(EffectSizes, align = 'c', digits = 3, caption = "Partial Eta-Squared Effect Sizes")
```

### Two-way comparisons

```{r}
# Cell means for Bio at each level of Diet
d = describeBy(MAP ~ Diet + Bio, data = A4Data, mat = TRUE, digits = 2)
names(d)[names(d) == "group1"] = "Diet"
names(d)[names(d) == "group2"] = "Bio"
rownames(d) = NULL
d[c(2,3,5:7,16)] %>%
  kable(caption = "Descriptives: Bio at each level of Diet", align = 'c', digits = 3)
```

```{r}
emmeans(A4LM, pairwise ~ Bio | Diet, adjust = AdjustmentChoice)$contrasts %>% 
  kable(caption = "Pairwise Bio at each level of Diet", align = 'c', digits = 3)
```

```{r}
# Cell means for Bio at each level of Drug
d = describeBy(MAP ~ Drug + Bio, data = A4Data, mat = TRUE, digits = 2)
names(d)[names(d) == "group1"] = "Drug"
names(d)[names(d) == "group2"] = "Bio"
rownames(d) = NULL
d[c(2,3,5:7,16)] %>%
  kable(caption = "Descriptives: Bio at each level of Drug", align = 'c', digits = 3)
```

```{r}
emmeans(A4LM, pairwise ~ Bio | Drug, adjust = AdjustmentChoice)$contrasts %>% 
  kable(caption = "Pairwise Bio at each level of Drug", align = 'c', digits = 3)
```

```{r}
# Cell means for Diet at each level of Drug
d = describeBy(MAP ~ Drug + Diet, data = A4Data, mat = TRUE, digits = 2)
names(d)[names(d) == "group1"] = "Drug"
names(d)[names(d) == "group2"] = "Diet"
rownames(d) = NULL
d[c(2,3,5:7,16)] %>%
  kable(caption = "Descriptives: Diet at each level of Drug", align = 'c', digits = 3)
```

```{r}
emmeans(A4LM, pairwise ~ Diet | Drug, adjust = AdjustmentChoice)$contrasts %>% 
  kable(caption = "Pairwise Diet at each level of Drug", align = 'c', digits = 3)
```

```{r}
# Cell means for Diet at each level of Bio
d = describeBy(MAP ~ Bio + Diet, data = A4Data, mat = TRUE, digits = 2)
names(d)[names(d) == "group1"] = "Bio"
names(d)[names(d) == "group2"] = "Diet"
rownames(d) = NULL
d[c(2,3,5:7,16)] %>%
  kable(caption = "Descriptives: Diet at each level of Bio", align = 'c', digits = 3)
```

```{r}
emmeans(A4LM, pairwise ~ Diet | Bio, adjust = AdjustmentChoice)$contrasts %>% 
  kable(caption = "Pairwise Diet at each level of Bio", align = 'c', digits = 3)
```

```{r}
# Cell means for Drug at each level of Diet
d = describeBy(MAP ~ Diet + Drug, data = A4Data, mat = TRUE, digits = 2)
names(d)[names(d) == "group1"] = "Diet"
names(d)[names(d) == "group2"] = "Drug"
rownames(d) = NULL
d[c(2,3,5:7,16)] %>%
  kable(caption = "Descriptives: Drug at each level of Diet", align = 'c', digits = 3)
```

### To justify t-tests within variables which have 3 or more levels you need a successful f-test.

```{r}
testInteractions(A4LM, 
                 fixed  = "Diet", 
                 across = "Drug") %>% kable(align = 'c', digits = 3)
```

```{r}
emmeans(A4LM, pairwise ~ Drug | Diet, adjust = AdjustmentChoice)$contrasts %>% 
  kable(caption = "Pairwise Drug at each level of Diet", align = 'c', digits = 3)
```

```{r}
# Cell means for Drug at each level of Bio
d = describeBy(MAP ~ Bio + Drug, data = A4Data, mat = TRUE, digits = 2)
names(d)[names(d) == "group1"] = "Bio"
names(d)[names(d) == "group2"] = "Drug"
rownames(d) = NULL
d[c(2,3,5:7,16)] %>%
  kable(caption = "Descriptives: Drug at each level of Bio", align = 'c', digits = 3)
```

### Again, to justify t-tests within variables which have 3 or more levels you need a successful f-test.

```{r}
testInteractions(A4LM, 
                 fixed  = "Bio", 
                 across = "Drug") %>% kable(align = 'c', digits = 3)
```

```{r}
emmeans(A4LM, pairwise ~ Drug | Bio, adjust = AdjustmentChoice)$contrasts %>% 
  kable(caption = "Pairwise Drug at each level of Bio", align = 'c', digits = 3)
```


### Three-way comparisons

```{r}
# Cell means for Diet at each Bio × Drug combination
d = describeBy(MAP ~ Bio + Drug + Diet, data = A4Data, mat = TRUE, digits = 2)
names(d)[names(d) == "group1"] = "Bio"
names(d)[names(d) == "group2"] = "Drug"
names(d)[names(d) == "group3"] = "Diet"
rownames(d) = NULL
d[c(2,3,4,6:8,17)] %>%
  kable(caption = "Descriptives: Diet at each Bio × Drug combination", align = 'c', digits = 3)
```

```{r}
emmeans(A4LM, pairwise ~ Diet | Bio * Drug, adjust = AdjustmentChoice)$contrasts %>%
  kable(caption = "Pairwise Diet at each Bio × Drug combination", align = 'c', digits = 3)
```

```{r}
# Cell means for Bio at each Diet × Drug combination
d = describeBy(MAP ~ Diet + Drug + Bio, data = A4Data, mat = TRUE, digits = 2)
names(d)[names(d) == "group1"] = "Diet"
names(d)[names(d) == "group2"] = "Drug"
names(d)[names(d) == "group3"] = "Bio"
rownames(d) = NULL
d[c(2,3,4,6:8,17)] %>%
  kable(caption = "Descriptives: Bio at each Diet × Drug combination", align = 'c', digits = 3)
```

```{r}
emmeans(A4LM, pairwise ~ Bio | Diet * Drug, adjust = AdjustmentChoice)$contrasts %>%
  kable(caption = "Pairwise Bio at each Diet × Drug combination", align = 'c', digits = 3)
```

```{r}
# Cell means for Drug at each Diet × Bio combination
d = describeBy(MAP ~ Diet + Bio + Drug, data = A4Data, mat = TRUE, digits = 2)
names(d)[names(d) == "group1"] = "Diet"
names(d)[names(d) == "group2"] = "Bio"
names(d)[names(d) == "group3"] = "Drug"
rownames(d) = NULL
d[c(2,3,4,6:8,17)] %>%
  kable(caption = "Descriptives: Drug at each Diet × Bio combination", align = 'c', digits = 3)
```

### Again, to justify t-tests within variables which have 3 or more levels you need a successful f-test. In the first column, the first number is the Diet level the second is the Bio level.

```{r}
testInteractions(A4LM,
                 fixed = c("Diet", "Bio"),
                 across = "Drug") %>% 
  kable(align = 'c', digits = 3)
```

```{r}
emmeans(A4LM, pairwise ~ Drug | Diet * Bio, adjust = AdjustmentChoice)$contrasts %>%
  kable(caption = "Pairwise Drug at each Diet × Bio combination", align = 'c', digits = 3)
```
