---
title: "Interaction supplement"
format: docx
execute:
   echo: false
   warning: false
   fig-dpi: 300
---

```{r}
library(tidyverse)
library(knitr)
library(car)
library(psych)
library(emmeans)
library(phia)
library(interactions)
```

---

### Main effect of A
```{r}
Scores = data.frame(B1 = c(10,20),
                    B2 = c(10,20))

rownames(Scores) = c("A1","A2")

ymax = max(Scores) + 3
ymin = min(Scores) - 3

xmax = ncol(Scores) + 1

plot(Scores$B1,
     xlim = c(0,xmax),
     ylim = c(ymin,ymax),
     type = 'both',
     pch = 16,
     ylab = "Score",
     xlab = '',
     xaxt = 'n',
     lwd = 3,
     col = 'blue')

lines(Scores$B2,
      type = 'both',
      pch = 16,
      lw = 3,
      col = 'darkgreen')

axis(1, at = 1:2, labels = rownames(Scores))

legend("topleft",
       legend = c("B1", "B2"),
       col    = c("blue", "darkgreen"),
       lwd    = 3,
       pch    = c(16, 17),
       bty    = "n")            # no box around the legend

kable(Scores, align = 'c')

```

---

### Main effect of B
```{r}
Scores = data.frame(B1 = c(10,10),
                    B2 = c(20,20))

rownames(Scores) = c("A1","A2")

ymax = max(Scores) + 3
ymin = min(Scores) - 3

xmax = ncol(Scores) + 1

plot(Scores$B1,
     xlim = c(0,xmax),
     ylim = c(ymin,ymax),
     type = 'both',
     pch = 16,
     ylab = "Score",
     xlab = '',
     xaxt = 'n',
     lwd = 3,
     col = 'blue')

lines(Scores$B2,
      type = 'both',
      pch = 16,
      lw = 3,
      col = 'darkgreen')

axis(1, at = 1:2, labels = rownames(Scores))

legend("topleft",
       legend = c("B1", "B2"),
       col    = c("blue", "darkgreen"),
       lwd    = 3,
       pch    = c(16, 17),
       bty    = "n")            # no box around the legend

kable(Scores, align = 'c')

```

---

### Main effect of A & B (no interaction)
```{r}
Scores = data.frame(B1 = c(10,20),
                    B2 = c(20,30))

rownames(Scores) = c("A1","A2")

ymax = max(Scores) + 3
ymin = min(Scores) - 3

xmax = ncol(Scores) + 1

plot(Scores$B1,
     xlim = c(0,xmax),
     ylim = c(ymin,ymax),
     type = 'both',
     pch = 16,
     ylab = "Score",
     xlab = '',
     xaxt = 'n',
     lwd = 3,
     col = 'blue')

lines(Scores$B2,
      type = 'both',
      pch = 16,
      lw = 3,
      col = 'darkgreen')

axis(1, at = 1:2, labels = rownames(Scores))

legend("topleft",
       legend = c("B1", "B2"),
       col    = c("blue", "darkgreen"),
       lwd    = 3,
       pch    = c(16, 17),
       bty    = "n")            # no box around the legend

kable(Scores, align = 'c')

```

---

### No effect
```{r}
Scores = data.frame(B1 = c(15,15),
                    B2 = c(15,15))

rownames(Scores) = c("A1","A2")

ymax = max(Scores) + 3
ymin = min(Scores) - 3

xmax = ncol(Scores) + 1

plot(Scores$B1,
     xlim = c(0,xmax),
     ylim = c(ymin,ymax),
     type = 'both',
     pch = 16,
     ylab = "Score",
     xlab = '',
     xaxt = 'n',
     lwd = 3,
     col = 'blue')

lines(Scores$B2,
      type = 'both',
      pch = 16,
      lw = 3,
      col = 'darkgreen')

axis(1, at = 1:2, labels = rownames(Scores))

legend("topleft",
       legend = c("B1", "B2"),
       col    = c("blue", "darkgreen"),
       lwd    = 3,
       pch    = c(16, 17),
       bty    = "n")            # no box around the legend

kable(Scores, align = 'c')

```

---

### One-way interaction with no main effect
```{r}
Scores = data.frame(B1 = c(10,20),
                    B2 = c(20,10))

rownames(Scores) = c("A1","A2")

ymax = max(Scores) + 3
ymin = min(Scores) - 3

xmax = ncol(Scores) + 1

plot(Scores$B1,
     xlim = c(0,xmax),
     ylim = c(ymin,ymax),
     type = 'both',
     pch = 16,
     ylab = "Score",
     xlab = '',
     xaxt = 'n',
     lwd = 3,
     col = 'blue')

lines(Scores$B2,
      type = 'both',
      pch = 16,
      lw = 3,
      col = 'darkgreen')

axis(1, at = 1:2, labels = rownames(Scores))

legend("topleft",
       legend = c("B1", "B2"),
       col    = c("blue", "darkgreen"),
       lwd    = 3,
       pch    = c(16, 17),
       bty    = "n")            # no box around the legend

kable(Scores, align = 'c')

```

---

### One-way interaction with main effects of both A & B
```{r}
Scores = data.frame(B1 = c(10,20),
                    B2 = c(20,40))

rownames(Scores) = c("A1","A2")

ymax = max(Scores) + 3
ymin = min(Scores) - 3

xmax = ncol(Scores) + 1

plot(Scores$B1,
     xlim = c(0,xmax),
     ylim = c(ymin,ymax),
     type = 'both',
     pch = 16,
     ylab = "Score",
     xlab = '',
     xaxt = 'n',
     lwd = 3,
     col = 'blue')

lines(Scores$B2,
      type = 'both',
      pch = 16,
      lw = 3,
      col = 'darkgreen')

axis(1, at = 1:2, labels = rownames(Scores))

legend("topleft",
       legend = c("B1", "B2"),
       col    = c("blue", "darkgreen"),
       lwd    = 3,
       pch    = c(16, 17),
       bty    = "n")            # no box around the legend

kable(Scores, align = 'c')

```

---

### Another one-way interaction with main effects of both A & B
```{r}
Scores = data.frame(B1 = c(40,20),
                    B2 = c(10,20))

rownames(Scores) = c("A1","A2")

ymax = max(Scores) + 3
ymin = min(Scores) - 3

xmax = ncol(Scores) + 1

plot(Scores$B1,
     xlim = c(0,xmax),
     ylim = c(ymin,ymax),
     type = 'both',
     pch = 16,
     ylab = "Score",
     xlab = '',
     xaxt = 'n',
     lwd = 3,
     col = 'blue')

lines(Scores$B2,
      type = 'both',
      pch = 16,
      lw = 3,
      col = 'darkgreen')

axis(1, at = 1:2, labels = rownames(Scores))

legend("topleft",
       legend = c("B1", "B2"),
       col    = c("blue", "darkgreen"),
       lwd    = 3,
       pch    = c(16, 17),
       bty    = "n")            # no box around the legend

kable(Scores, align = 'c')

```

---

### Make your own Two-way
```{r}
####
#Change the scores to get a different effect
###
Scores = data.frame(B1 = c(40,20),
                    B2 = c(10,20))

### Don't make changes below here.

rownames(Scores) = c("A1","A2")

ymax = max(Scores) + 3
ymin = min(Scores) - 3

xmax = ncol(Scores) + 1

plot(Scores$B1,
     xlim = c(0,xmax),
     ylim = c(ymin,ymax),
     type = 'both',
     pch = 16,
     ylab = "Score",
     xlab = '',
     xaxt = 'n',
     lwd = 3,
     col = 'blue')

lines(Scores$B2,
      type = 'both',
      pch = 16,
      lw = 3,
      col = 'darkgreen')

axis(1, at = 1:2, labels = rownames(Scores))

legend("topleft",
       legend = c("B1", "B2"),
       col    = c("blue", "darkgreen"),
       lwd    = 3,
       pch    = c(16, 17),
       bty    = "n")            # no box around the legend

kable(Scores, align = 'c')

```

---

### Three way interaction.
```{r}

Scores_C1 = data.frame(
  B1 = c(40, 20),
  B2 = c(10, 20)
)
rownames(Scores_C1) = c("A1", "A2")

Scores_C2 = data.frame(
  B1 = c(15, 35),
  B2 = c(30, 10)
)
rownames(Scores_C2) = c("A1", "A2")

ymax = max(c(as.matrix(Scores_C1), as.matrix(Scores_C2))) + 5
ymin = min(c(as.matrix(Scores_C1), as.matrix(Scores_C2))) - 5

# Margins and declaration for two panels
par(mfrow = c(1, 2), mar = c(5, 4, 3, 1))

#  Panel 1: C1
plot(Scores_C1$B1,
     xlim = c(0.5, 2.5),
     ylim = c(ymin, ymax),
     type = "b",
     pch  = 16,
     lwd  = 3,
     col  = "blue",
     xaxt = "n",
     xlab = "",
     ylab = "Score",
     main = "C1")
lines(Scores_C1$B2, type = "b", pch = 17, lwd = 3, col = "darkgreen")
axis(1, at = 1:2, labels = c("A1", "A2"))
legend("topleft", legend = c("B1", "B2"),
       col = c("blue", "darkgreen"), lwd = 3, pch = c(16, 17), bty = "n")

# Panel 2: C2 
plot(Scores_C2$B1,
     xlim = c(0.5, 2.5),
     ylim = c(ymin, ymax),
     type = "b",
     pch  = 16,
     lwd  = 3,
     col  = "blue",
     xaxt = "n",
     xlab = "",
     ylab = "",
     main = "C2")
lines(Scores_C2$B2, type = "b", pch = 17, lwd = 3, col = "darkgreen")
axis(1, at = 1:2, labels = c("A1", "A2"))
legend("topleft", legend = c("B1", "B2"),
       col = c("blue", "darkgreen"), lwd = 3, pch = c(16, 17), bty = "n")

# Reset the plotting layout
par(mfrow = c(1, 1))

```

---

### Two way interaction with three variables, no main effect of C
```{r}
Scores_C1 = data.frame(
  B1 = c(40, 40),
  B2 = c(40, 20)
)
rownames(Scores_C1) = c("A1", "A2")

Scores_C2 = data.frame(
  B1 = c(40, 40),
  B2 = c(40, 20)
)
rownames(Scores_C2) = c("A1", "A2")

ymax = max(c(as.matrix(Scores_C1), as.matrix(Scores_C2))) + 5
ymin = min(c(as.matrix(Scores_C1), as.matrix(Scores_C2))) - 5

# Margins and declaration for two panels
par(mfrow = c(1, 2), mar = c(5, 4, 3, 1))

#  Panel 1: C1
plot(Scores_C1$B1,
     xlim = c(0.5, 2.5),
     ylim = c(ymin, ymax),
     type = "b",
     pch  = 16,
     lwd  = 3,
     col  = "blue",
     xaxt = "n",
     xlab = "",
     ylab = "Score",
     main = "C1")
lines(Scores_C1$B2, type = "b", pch = 17, lwd = 3, col = "darkgreen")
axis(1, at = 1:2, labels = c("A1", "A2"))
legend("topleft", legend = c("B1", "B2"),
       col = c("blue", "darkgreen"), lwd = 3, pch = c(16, 17), bty = "n")

# Panel 2: C2 
plot(Scores_C2$B1,
     xlim = c(0.5, 2.5),
     ylim = c(ymin, ymax),
     type = "b",
     pch  = 16,
     lwd  = 3,
     col  = "blue",
     xaxt = "n",
     xlab = "",
     ylab = "",
     main = "C2")
lines(Scores_C2$B2, type = "b", pch = 17, lwd = 3, col = "darkgreen")
axis(1, at = 1:2, labels = c("A1", "A2"))
legend("topleft", legend = c("B1", "B2"),
       col = c("blue", "darkgreen"), lwd = 3, pch = c(16, 17), bty = "n")

# Reset the plotting layout
par(mfrow = c(1, 1))

```

---

### Two way interaction with three variables, with a main effect of C
```{r}
Scores_C1 = data.frame(
  B1 = c(40, 40),
  B2 = c(40, 20)
)
rownames(Scores_C1) = c("A1", "A2")

Scores_C2 = data.frame(
  B1 = c(30, 30),
  B2 = c(30, 10)
)
rownames(Scores_C2) = c("A1", "A2")

ymax = max(c(as.matrix(Scores_C1), as.matrix(Scores_C2))) + 5
ymin = min(c(as.matrix(Scores_C1), as.matrix(Scores_C2))) - 5

# Margins and declaration for two panels
par(mfrow = c(1, 2), mar = c(5, 4, 3, 1))

#  Panel 1: C1
plot(Scores_C1$B1,
     xlim = c(0.5, 2.5),
     ylim = c(ymin, ymax),
     type = "b",
     pch  = 16,
     lwd  = 3,
     col  = "blue",
     xaxt = "n",
     xlab = "",
     ylab = "Score",
     main = "C1")
lines(Scores_C1$B2, type = "b", pch = 17, lwd = 3, col = "darkgreen")
axis(1, at = 1:2, labels = c("A1", "A2"))
legend("topleft", legend = c("B1", "B2"),
       col = c("blue", "darkgreen"), lwd = 3, pch = c(16, 17), bty = "n")

# Panel 2: C2 
plot(Scores_C2$B1,
     xlim = c(0.5, 2.5),
     ylim = c(ymin, ymax),
     type = "b",
     pch  = 16,
     lwd  = 3,
     col  = "blue",
     xaxt = "n",
     xlab = "",
     ylab = "",
     main = "C2")
lines(Scores_C2$B2, type = "b", pch = 17, lwd = 3, col = "darkgreen")
axis(1, at = 1:2, labels = c("A1", "A2"))
legend("topleft", legend = c("B1", "B2"),
       col = c("blue", "darkgreen"), lwd = 3, pch = c(16, 17), bty = "n")

# Reset the plotting layout
par(mfrow = c(1, 1))

```


---

### Make your own.
```{r}
####
#Change the scores to get a different effect
###

Scores_C1 = data.frame(
  B1 = c(40, 40),
  B2 = c(40, 20)
)
rownames(Scores_C1) = c("A1", "A2")

Scores_C2 = data.frame(
  B1 = c(30, 30),
  B2 = c(30, 20)
)
rownames(Scores_C2) = c("A1", "A2")


### Don't make any changes below
ymax = max(c(as.matrix(Scores_C1), as.matrix(Scores_C2))) + 5
ymin = min(c(as.matrix(Scores_C1), as.matrix(Scores_C2))) - 5

# Margins and declaration for two panels
par(mfrow = c(1, 2), mar = c(5, 4, 3, 1))

#  Panel 1: C1
plot(Scores_C1$B1,
     xlim = c(0.5, 2.5),
     ylim = c(ymin, ymax),
     type = "b",
     pch  = 16,
     lwd  = 3,
     col  = "blue",
     xaxt = "n",
     xlab = "",
     ylab = "Score",
     main = "C1")
lines(Scores_C1$B2, type = "b", pch = 17, lwd = 3, col = "darkgreen")
axis(1, at = 1:2, labels = c("A1", "A2"))
legend("topleft", legend = c("B1", "B2"),
       col = c("blue", "darkgreen"), lwd = 3, pch = c(16, 17), bty = "n")

# Panel 2: C2 
plot(Scores_C2$B1,
     xlim = c(0.5, 2.5),
     ylim = c(ymin, ymax),
     type = "b",
     pch  = 16,
     lwd  = 3,
     col  = "blue",
     xaxt = "n",
     xlab = "",
     ylab = "",
     main = "C2")
lines(Scores_C2$B2, type = "b", pch = 17, lwd = 3, col = "darkgreen")
axis(1, at = 1:2, labels = c("A1", "A2"))
legend("topleft", legend = c("B1", "B2"),
       col = c("blue", "darkgreen"), lwd = 3, pch = c(16, 17), bty = "n")

# Reset the plotting layout
par(mfrow = c(1, 1))

```