---
title: "Lab 9"
format: html
execute:
   echo: false
   warning: false
   fig-dpi: 300
---


```{r}
library(tidyverse)
library(knitr)
library(car)
library(psych)
library(emmeans)
library(phia)
library(interactions)
```


```{r}
# Load your data
ReportingExampleData = read.table("3-way ANOVA Lecture.csv",
                          sep = ",",
                          header = T,
                          colClasses = c("factor",
                                         "factor",
                                         "factor",
                                         "integer"))
```

### Your data
Participants reported whether they had children (Kids: 1 = no, 2 = yes), whether they had voted in the most recent election (Vote: 1 = no, 2 = yes), and their political orientation (Politics: 1 = liberal, 2 = moderate, 3 = conservative). The outcome variable, Score, was a continuous measure of participants’ attitudes toward a proposed education policy.
```{r}
kable (ReportingExampleData,align = 'c', digits = 3)

vars = c("Kids", "Vote", "Politics")
a = NULL
for (v in vars) {
   for (lev in levels(ReportingExampleData[[v]])) {
      desc = describe(ReportingExampleData[ReportingExampleData[[v]] == lev, "Score"])[c(2:4)]
      a = rbind(a, desc)
   }
}
rownames(a) = c(
   paste0("Kids", levels(ReportingExampleData$Kids)),
   paste0("Vote", levels(ReportingExampleData$Vote)),
   paste0("Politics", levels(ReportingExampleData$Politics))
)

kable(a, align = 'c', digits = 3)

```


```{r}

# "contr.sum" is the correct option for typle type III SS
options(contrasts = c("contr.sum", "contr.poly"))
# "contr.treatment" is a reference group type contrast closer to a type I or
# type II SS.
#options(contrasts = c("contr.treatment", "contr.poly"))

leveneTest(Score ~ Kids * Vote * Politics,ReportingExampleData)
```

### Linear Model
```{r}
ReportingExLM = lm(Score ~ Kids * Vote * Politics, ReportingExampleData )

summary(ReportingExLM)

```

# Omnibus
```{r}
a = Anova(ReportingExLM, type = 3)

SS_error = a["Residuals", "Sum Sq"]

a$pes = a$`Sum Sq` / (a$`Sum Sq` + SS_error)
a["Residuals", "pes"] = NA          
a["(Intercept)", "pes"] = NA 
kable(a, align = 'c', digits = 3)
```

```{r}
# Post hoc adjustment type. Uncomment only one choice.
AdjustmentChoice = "none" # Least significant difference
# AdjustmentChoice = "bonferrroni"
# AdjustmentChoice = "tukey"
```

### Two-way comparisons

```{r}
# Cell means for Vote at each level of Kids
d = describeBy(Score ~ Kids + Vote, data = ReportingExampleData, mat = TRUE, digits = 2)
names(d)[names(d) == "group1"] = "Kids"
names(d)[names(d) == "group2"] = "Vote"
rownames(d) = NULL
d[c(2,3,5:7,16)] %>%
  kable(caption = "Descriptives: Vote at each level of Kids", align = 'c', digits = 3)
```

```{r}
emmeans(ReportingExLM, pairwise ~ Vote | Kids, adjust = AdjustmentChoice)$contrasts %>% 
  kable(caption = "Pairwise Vote at each level of Kids", align = 'c', digits = 3)
```

```{r}
# Cell means for Vote at each level of Politics
d = describeBy(Score ~ Politics + Vote, data = ReportingExampleData, mat = TRUE, digits = 2)
names(d)[names(d) == "group1"] = "Politics"
names(d)[names(d) == "group2"] = "Vote"
rownames(d) = NULL
d[c(2,3,5:7,16)] %>%
  kable(caption = "Descriptives: Vote at each level of Politics", align = 'c', digits = 3)
```

```{r}
emmeans(ReportingExLM, pairwise ~ Vote | Politics, adjust = AdjustmentChoice)$contrasts %>% 
  kable(caption = "Pairwise Vote at each level of Politics", align = 'c', digits = 3)
```

```{r}
# Cell means for Kids at each level of Politics
d = describeBy(Score ~ Politics + Kids, data = ReportingExampleData, mat = TRUE, digits = 2)
names(d)[names(d) == "group1"] = "Politics"
names(d)[names(d) == "group2"] = "Kids"
rownames(d) = NULL
d[c(2,3,5:7,16)] %>%
  kable(caption = "Descriptives: Kids at each level of Politics", align = 'c', digits = 3)
```

```{r}
emmeans(ReportingExLM, pairwise ~ Kids | Politics, adjust = AdjustmentChoice)$contrasts %>% 
  kable(caption = "Pairwise Kids at each level of Politics", align = 'c', digits = 3)
```

```{r}
# Cell means for Kids at each level of Vote
d = describeBy(Score ~ Vote + Kids, data = ReportingExampleData, mat = TRUE, digits = 2)
names(d)[names(d) == "group1"] = "Vote"
names(d)[names(d) == "group2"] = "Kids"
rownames(d) = NULL
d[c(2,3,5:7,16)] %>%
  kable(caption = "Descriptives: Kids at each level of Vote", align = 'c', digits = 3)
```

```{r}
emmeans(ReportingExLM, pairwise ~ Kids | Vote, adjust = AdjustmentChoice)$contrasts %>% 
  kable(caption = "Pairwise Kids at each level of Vote", align = 'c', digits = 3)
```

```{r}
# Cell means for Politics at each level of Kids
d = describeBy(Score ~ Kids + Politics, data = ReportingExampleData, mat = TRUE, digits = 2)
names(d)[names(d) == "group1"] = "Kids"
names(d)[names(d) == "group2"] = "Politics"
rownames(d) = NULL
d[c(2,3,5:7,16)] %>%
  kable(caption = "Descriptives: Politics at each level of Kids", align = 'c', digits = 3)
```

### To justify t-tests within variables which have 3 or more levels you need a successful f-test.

```{r}
testInteractions(ReportingExLM, 
                 fixed  = "Kids", 
                 across = "Politics") %>% kable(align = 'c', digits = 3)
```

### The first level of kids was successful, so you are justified in doing deeper testng. You can use the t-tests for the different levels of politics within level 1 of kids, but not level 2.

```{r}
emmeans(ReportingExLM, pairwise ~ Politics | Kids, adjust = AdjustmentChoice)$contrasts %>% 
  kable(caption = "Pairwise Kids at each level of Politics", align = 'c', digits = 3)
```

```{r}
# Cell means for Politics at each level of Vote
d = describeBy(Score ~ Vote + Politics, data = ReportingExampleData, mat = TRUE, digits = 2)
names(d)[names(d) == "group1"] = "Vote"
names(d)[names(d) == "group2"] = "Politics"
rownames(d) = NULL
d[c(2,3,5:7,16)] %>%
  kable(caption = "Descriptives: Politics at each level of Vote", align = 'c', digits = 3)
```

### Again, to justify t-tests within variables which have 3 or more levels you need a successful f-test.

```{r}
testInteractions(ReportingExLM, 
                 fixed  = "Vote", 
                 across = "Politics") %>% kable(align = 'c', digits = 3)
```

### Here, both levels of vote have significant findings so you should report all the signifiant t-tests below.

```{r}
emmeans(ReportingExLM, pairwise ~ Politics | Vote, adjust = AdjustmentChoice)$contrasts %>% 
  kable(caption = "Pairwise Politics at each level of Vote", align = 'c', digits = 3)
```


### Three-way comparisons

```{r}
# Cell means for Kids at each Vote × Politics combination
d = describeBy(Score ~ Vote + Politics + Kids, data = ReportingExampleData, mat = TRUE, digits = 2)
names(d)[names(d) == "group1"] = "Vote"
names(d)[names(d) == "group2"] = "Politics"
names(d)[names(d) == "group3"] = "Kids"
rownames(d) = NULL
d[c(2,3,4,6:8,17)] %>%
  kable(caption = "Descriptives: Kids at each Vote × Politics combination", align = 'c', digits = 3)
```

```{r}
emmeans(ReportingExLM, pairwise ~ Kids | Vote * Politics, adjust = AdjustmentChoice)$contrasts %>%
  kable(caption = "Pairwise Kids at each Vote × Politics combination", align = 'c', digits = 3)
```

```{r}
# Cell means for Vote at each Kids × Politics combination
d = describeBy(Score ~ Kids + Politics + Vote, data = ReportingExampleData, mat = TRUE, digits = 2)
names(d)[names(d) == "group1"] = "Kids"
names(d)[names(d) == "group2"] = "Politics"
names(d)[names(d) == "group3"] = "Vote"
rownames(d) = NULL
d[c(2,3,4,6:8,17)] %>%
  kable(caption = "Descriptives: Vote at each Kids × Politics combination", align = 'c', digits = 3)
```

```{r}
emmeans(ReportingExLM, pairwise ~ Vote | Kids * Politics, adjust = "none")$contrasts %>%
  kable(caption = "Pairwise Vote at each Kids × Politics combination", align = 'c', digits = 3)
```

```{r}
# Cell means for Politics at each Kids × Vote combination
d = describeBy(Score ~ Kids + Vote + Politics, data = ReportingExampleData, mat = TRUE, digits = 2)
names(d)[names(d) == "group1"] = "Kids"
names(d)[names(d) == "group2"] = "Vote"
names(d)[names(d) == "group3"] = "Politics"
rownames(d) = NULL
d[c(2,3,4,6:8,17)] %>%
  kable(caption = "Descriptives: Politics at each Kids × Vote combination", align = 'c', digits = 3)
```

### Again, to justify t-tests within variables which have 3 or more levels you need a successful f-test. In the first column, the first number is the Kids level the second is the Vote level.

```{r}
testInteractions(ReportingExLM,
                 fixed = c("Kids", "Vote"),
                 across = "Politics") %>% 
  kable(align = 'c', digits = 3)
```

### All the tests were significant. So, if you choose to report this section, you should report all of the t-tests.

```{r}
emmeans(ReportingExLM, pairwise ~ Politics | Kids * Vote, adjust = "none")$contrasts %>%
  kable(caption = "Pairwise Politics at each Kids × Vote combination", align = 'c', digits = 3)
```
