---
title: "Assignment 2"
format: html
execute:
   echo: FALSE
   warning: FALSE
   message: FALSE
---

```{r}
library(knitr)
library(psych)      # describe(), describeBy()
library(car)        # leveneTest() 
library(effectsize) # cohens_d()
```


---

### Submission notes

- Render this document into a word document, input your answers, and submit the finished word document together with this `.qmd` source.

- Do not put your name anywhere in the document (anonymous grading).

- All of your responses to questions 1-5 should be manually input to the Word document you generate here. All of you numerical results for questions 6-9 should come up when the grader runs this script. However, you still need to manually write the specific statistics and  the short interpretations which are requested below.


### Overview & Instructions for questions four through six.

Your main task is to choose the correct test for the given research context and report the following items.

a. Is the test significant? (α = .05, two-tailed). Include the t value, df, and p-value from the output.

b. Cohen’s d and briefly interpret its magnitude (small / medium / large).

c. Confidence interval- Report the 95 % CI for the mean difference. Does the interval contain 0? What does that tell you?

The code deliberately lets you run different statistical processes by setting logical flags and uncommenting code. Setting the flags incorrectly or choosing the wrong code produces a statistically legal but conceptually wrong analysis — exactly the kind of mistake that is easy to make when the design is not carefully considered.

- Read the research context carefully.
- Decide which testing method to use.
- Uncomment code and set the flag(s) (i.e., `paired_flag = T` or `F`) accordingly.
- Render this script into a Word document, then produce the items listed above (a through c).

You may also change the outcome variable (column name) if you want to explore but be sure to restore the original variable because that is what will be graded.


---

### 4. Transmission type and fuel efficiency  
Research context:
A researcher wants to know whether cars with automatic transmissions (`am == 0`) and cars with manual transmissions (`am == 1`) differ in fuel efficiency. The data come from the `mtcars` data set; each row is a different car. We will examine the `mpg` column.

If you wish to explore, you may change "mpg" to another continuous variable ("hp", "wt", "qsec", "disp",etc.). Remember though, the assignment is to examine mpg.

```{r}
# Creating groups and making them equal sizes.
group1 = mtcars[mtcars$am == 0, c("am","mpg")]   # automatic 
group2 = mtcars[mtcars$am == 1, c("am","mpg")]   # manual    
group1 = group1[7:19,]
AMDF = rbind(group1, group2)

# Descriptives
describe(group1$mpg)
describe(group2$mpg)
```

```{r}
# Optional:  Levene's test
# Uncomment the next line to run it.
# leveneTest(mpg ~ factor(am), AMDF)
```

```{r}
# Is this a paired / repeated-measures design or an independent-groups design?
# Set the flag to the value that matches the research context and your data.
paired_flag = F
Welch_flag = F

# Running and reporting the t-test you have chosen.
tt = t.test(group1$mpg, group2$mpg, paired = paired_flag, var.equal = !Welch_flag)
tt

# Cohen's d uses the same paired flag
d = cohens_d(group1$mpg, group2$mpg, paired = paired_flag)
d
```


---

### 5. Two soporific drugs and extra sleep  

The `sleep` data set records the extra hours of sleep obtained by subjects under two different drugs (`group`). Importantly, the same ten subjects each received both drugs (the rows are ordered by subject ID). The question is whether the two drugs produce different amounts of extra sleep.

```{r}
# Explicit vectors (order follows subject ID, so pairing is preserved)
group1 = sleep$extra[sleep$group == 1]
group2 = sleep$extra[sleep$group == 2]

# Descriptives
describe(group1)
describe(group2)
```

```{r}
# Optional:  Levene's test
# Uncomment the next line to run it.
# leveneTest(extra ~ factor(group), data = sleep)
```

```{r}
# Is this a paired / repeated-measures design or an independent-groups design?
# Set the flag to the value that matches the research context and your data.
paired_flag = F
Welch_flag = F

# Running and reporting the t-test you have chosen.
tt = t.test(group1, group2, paired = paired_flag, var.equal = !Welch_flag)
tt

# Cohen's d uses the same paired flag
d = cohens_d(group1, group2, paired = paired_flag)
d
```


---

### 6. Average fuel efficiency versus a benchmark  
Research context

Is the average fuel efficiency of the cars in `mtcars` different from a hypothesized value of 17 mpg? Use the code below to get your t and its p-value. Then, compare that p-value to an alpha of .05 to determine if you have found a significant difference.

```{r}
# Levene's test is not applicable to a one-sample design. (No action needed here.)
```

```{r}
# This will put your data into a variable called x. You can put x into a function to get information about your data.
x   = mtcars$mpg    

# This will give you all the descriptives about your data
describe(x)

# choose the functions below to get you the correct t-test formula. Comment them out so they don't appear in your Word document.
# for simplicity, one of the first four will give you your n.
n = 17
n = 17 - 1
n = length(mtcars$mpg)
n = length(mtcars$mpg) - 1
17 # Just the number 17
mean(17)
mean(mtcars$mpg)
var(17)
var(mtcars$mpg)
sd(17)
sd(mtcars$mpg)
mean(mtcars$mpg)
sqrt(17)
sqrt(mtcars$mpg)
sqrt(n)

# Put your chosen functions, values, or variables into the correct form to calculate your observed t.

numerator = 1

denominator = 2

t_score = numerator /  denominator

t_score

# Uncomment the next few lines to get your p-value.
# df = n - 1
# Tails = 2
# 
# pt(abs(t_score), df, lower.tail = F) * Tails


# Again, choose the correct functions, values, or variables for your numerator and denominator to get your Cohen's d.
 
d_numerator = 1

d_denominator = 2

cohens_d = abs(d_numerator / d_denominator)

cohens_d
```

