---
title: "t-test Practice Assignment"
subtitle: "Choosing and running the appropriate *t*-test"
author: "Student Name (leave blank for anonymous grading)"
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: false
    theme: cosmo
editor: source
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  comment = ""
)
library(psych)      # describe(), describeBy()
library(car)        # leveneTest() — optional
library(effectsize) # cohens_d()
```

## Overview & Instructions

This assignment rewrites the three calculation problems (original questions 4–6) using the familiar built-in datasets `mtcars` and `sleep`. The code style matches the preceding labs (explicit `group1` / `group2` subsets, base R + `psych` / `effectsize` / `car`).

**Your main task is to choose the correct test.**

For each problem a research context is given. The code deliberately lets you run *either* a paired or an independent-groups analysis by setting a single logical flag. Setting the flag incorrectly produces a statistically legal but conceptually wrong analysis — exactly the kind of mistake that is easy to make when the design is not carefully considered.

- Read the research context carefully.
- Decide whether the observations are paired / repeated-measures or independent.
- Set the flag (`paired_flag = TRUE` or `FALSE`) accordingly.
- Run the chunks, then answer a–c.

You may also change the outcome variable (column name) if you want to explore. Optional assumption checks remain commented; uncomment them if you wish.

No Excel or SPSS is required. Focus on correct test selection and clear interpretation of the R output.

---

## 1. Transmission type and fuel efficiency  
*(rewritten from original problem 4)*

**Research context**  
A researcher wants to know whether cars with automatic transmissions (`am == 0`) and cars with manual transmissions (`am == 1`) differ in fuel efficiency. The data come from the `mtcars` data set; each row is a different car. We will examine the `mpg` column.

You may change `"mpg"` to another continuous variable (`"hp"`, `"wt"`, `"qsec"`, `"disp"`, …) in the two places marked below.

```{r problem1-groups}
# Explicit group subsets
group1 = mtcars[mtcars$am == 0, "mpg"]   # automatic  — change "mpg" here if desired
group2 = mtcars[mtcars$am == 1, "mpg"]   # manual     — and here

# Descriptives
describe(group1)
describe(group2)
```

```{r problem1-levene}
# Optional: homogeneity of variance (Levene's test)
# Simply uncomment the next line to run it
# leveneTest(mpg ~ factor(am), data = mtcars)
```

```{r problem1-ttest}
# *** STUDENT CHOICE ***
# Is this a paired / repeated-measures design or an independent-groups design?
# Set the flag to the value that matches the research context.
# (Setting it incorrectly is a common error — think before you change it.)
paired_flag = FALSE

tt = t.test(group1, group2, paired = paired_flag)
tt

# Cohen's d uses the same flag
d = cohens_d(group1, group2, paired = paired_flag)
d
```

**a. Is the test significant?** (α = .05, two-tailed)  

*Write your answer here (include the *t* value, *df*, and *p*-value from the output).*

**b. Cohen’s *d***  

*Report the value and briefly interpret its magnitude (small / medium / large).*

**c. Confidence interval**  

*Report the 95 % CI for the mean difference. Does the interval contain 0? What does that tell you?*

---

## 2. Two soporific drugs and extra sleep  
*(rewritten from original problem 5)*

**Research context**  
The `sleep` data set records the *extra* hours of sleep obtained by subjects under two different drugs (`group`). Importantly, the *same ten subjects* each received both drugs (the rows are ordered by subject ID). The question is whether the two drugs produce different amounts of extra sleep.

```{r problem2-groups}
# Explicit vectors (order follows subject ID, so pairing is preserved)
group1 = sleep$extra[sleep$group == 1]
group2 = sleep$extra[sleep$group == 2]

# Descriptives by group
describeBy(sleep$extra, group = sleep$group)
```

```{r problem2-ttest}
# *** STUDENT CHOICE ***
# Is this a paired / repeated-measures design or an independent-groups design?
# Set the flag to the value that matches the research context.
# (Setting it incorrectly is a common error — think before you change it.)
paired_flag = TRUE

tt = t.test(group1, group2, paired = paired_flag)
tt

# Cohen's d uses the same flag
d = cohens_d(group1, group2, paired = paired_flag)
d
```

**a. Is the test significant?** (α = .05, two-tailed)  

*Write your answer here (include the *t* value, *df*, and *p*-value from the output).*

**b. Cohen’s *d***  

*Report the value and briefly interpret its magnitude (small / medium / large).*

**c. Confidence interval**  

*Report the 95 % CI for the mean difference. Does the interval contain 0? What does that tell you?*

---

## 3. Average fuel efficiency versus a benchmark  
*(rewritten from original problem 6)*

**Research context**  
Is the average fuel efficiency of the cars in `mtcars` different from a hypothesized value of 20 mpg? Only one sample of cars is available and the population standard deviation is unknown.

You may change the variable or the null value below if you wish to explore.

```{r problem3-prep}
# Choose the variable and the null hypothesis value
x   = mtcars$mpg    # change "mpg" if desired
mu0 = 20            # change the hypothesized mean if desired

describe(x)
```

```{r problem3-ttest}
# This problem compares one sample mean against a known constant.
# Do not attempt to turn it into a two-group (independent or paired) test.
# The correct call uses the mu argument.

tt = t.test(x, mu = mu0)
tt

# Cohen's d for a one-sample design
d = cohens_d(x, mu = mu0)
d
```

**a. Is the test significant?** (α = .05, two-tailed)  

*Write your answer here (include the *t* value, *df*, and *p*-value from the output).*

**b. Cohen’s *d***  

*Report the value and briefly interpret its magnitude (small / medium / large).*

**c. Confidence interval**  

*Report the 95 % CI for the mean. Does the interval contain the hypothesized value (μ₀)? What does that tell you?*

---

## Submission notes

- Knit this document to HTML (or PDF) and submit the rendered file together with the `.qmd` source.
- Do **not** put your name anywhere in the document (anonymous grading).
- All numerical results should come from the R output above; the written answers are short interpretations only.
- If you experimented with different variables or deliberately tried the wrong flag setting, briefly note what you did so the grader can reproduce your numbers.
```
