---
title: "Assignment 2"
format: html
execute:
   echo: false
   warning: false
   message: false
---

```{r}
library(knitr)
library(psych)      # describe(), describeBy()
library(car)        # leveneTest() 
library(effectsize) # cohens_d()
```

---

### Submission notes

- Render this document into a word document, input your answers, and submit the finished word document together with this `.qmd` source.

- Do not put your name anywhere in the document (anonymous grading).

- All numerical results should come show up when the grader runs the script above; the written answers are short interpretations only.


### Overview & Instructions for questions four through six.

Your main task is to choose the correct test for the given research context and report the following items.

a. Is the test significant? (α = .05, two-tailed). Include the t value, df, and p-value from the output.

b. Cohen’s d and briefly interpret its magnitude (small / medium / large).

c. Confidence interval- Report the 95 % CI for the mean difference. Does the interval contain 0? What does that tell you?

The code deliberately lets you run either a paired or an independent-groups analysis by setting a single logical flag. Setting the flag incorrectly produces a statistically legal but conceptually wrong analysis — exactly the kind of mistake that is easy to make when the design is not carefully considered.

- Read the research context carefully.
- Decide whether the observations are paired / repeated-measures or independent.
- Set the flag (`paired_flag = T` or `F`) accordingly.
- Run the chunks, then produce the items listed above (a through c).

You may also change the outcome variable (column name) if you want to explore.

Optional assumption checks remain commented; uncomment them if you wish.

---

### 4. Transmission type and fuel efficiency  
Research context:
A researcher wants to know whether cars with automatic transmissions (`am == 0`) and cars with manual transmissions (`am == 1`) differ in fuel efficiency. The data come from the `mtcars` data set; each row is a different car. We will examine the `mpg` column.

If you wish to explore, you may change "mpg" to another continuous variable ("hp", "wt", "qsec", "disp",etc.). Remeber though, the assignment is to  examine mpg.

```{r problem1-groups}
# Explicit group subsets
group1 = mtcars[mtcars$am == 0, "mpg"]   # automatic 
group2 = mtcars[mtcars$am == 1, "mpg"]   # manual    

# Descriptives
describe(group1)
describe(group2)
```

```{r problem1-levene}
# Optional: homogeneity of variance (Levene's test)
# Simply uncomment the next line to run it.
 leveneTest(mpg ~ factor(am), data = mtcars)
```

```{r problem1-ttest}
# Is this a paired / repeated-measures design or an independent-groups design?
# Set the flag to the value that matches the research context.
paired_flag = FALSE

tt = t.test(group1, group2, paired = paired_flag)
tt

# Cohen's d uses the same flag
d = cohens_d(group1, group2, paired = paired_flag)
d
```



---

### 5. Two soporific drugs and extra sleep  

The `sleep` data set records the extra hours of sleep obtained by subjects under two different drugs (`group`). Importantly, the same ten subjects each received both drugs (the rows are ordered by subject ID). The question is whether the two drugs produce different amounts of extra sleep.

```{r}
# Explicit vectors (order follows subject ID, so pairing is preserved)
group1 = sleep$extra[sleep$group == 1]
group2 = sleep$extra[sleep$group == 2]

# Descriptives by group
describeBy(sleep$extra, group = sleep$group)
```

```{r}
# Is this a paired / repeated-measures design or an independent-groups design?
# Set the flag to the value that matches the research context.

paired_flag = TRUE

tt = t.test(group1, group2, paired = paired_flag)
tt

# Cohen's d uses the same flag
d = cohens_d(group1, group2, paired = paired_flag)
d
```



---

### 6. Average fuel efficiency versus a benchmark  
Research context

Is the average fuel efficiency of the cars in `mtcars` different from a hypothesized value of 20 mpg? Only one sample of cars is available and the population standard deviation is unknown.

```{r}
# Choose the variable and the null hypothesis value
x   = mtcars$mpg    # change "mpg" if desired
mu0 = 20            # change the hypothesized mean if desired

describe(x)
```

```{r}


tt = t.test(x, mu = mu0)
tt

# Cohen's d for a one-sample design
d = cohens_d(x, mu = mu0)
d
```


